---
title: "R-Kurs Labb 1"
author: "Edvin Magnusson"
date: '2022-09-14'
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### R Markdown

### Uppgift 1  
Creating a vector function 


```{r}
# Clear the enviroment 
 rm(list = ls())

my_num_vector <- function(){
  x <- c(log10(11),cos(pi/5),exp(pi/3),(1173%%7)/19)
  return(x)
}

my_num_vector()

```

### Uppgift 2 
Create a function called mult_first_last() with the argument vector. The function shall return the
product of the first and last element in vector.

```{r}

mult_first_last<- function(vektor){
  
  result<- vektor[1]*tail(vektor,1)
  
  return(result)
  
}

mult_first_last(vektor = c(3,1,12,2,4))

mult_first_last(vektor = c(3,1,12))

```

### Uppgift 3 
Create a function called orth_scalar_prod() which calculate the scalar product between two vectors, a and b, in an orthonormal base. The scalar product is calculated in the following way:


```{r}

orth_scalar_prod<- function (a,b){
  
  result<- a%*%b
  return(result)
  
}

orth_scalar_prod(a = c(3,1,12,2,4), b = c(1,2,3,4,5))

orth_scalar_prod(a = c(-1, 3), b = c(-3, -1))

```

### Uppgift 4 


```{r}

lukes_father <- function(name){
  
  answer<- "I am your father."
  
  
  cat(name,answer,sep = ",")
  
}

lukes_father(name = "Luke")

lukes_father(name = "Edd")

```

### Uppgift 5 
Approximate the number e formula sum(1/n!)

```{r}
# The trick is that when n=0 and 1 the answer is 1+1 

approx_e<- function(N){
  
  i<- 2:N
  factorial_part <- 1/factorial(i)
  
  answer<- 1+1+sum(factorial_part)
  return(answer)

  
}

approx_e(N = 2)

approx_e(N = 4)

approx_e(N=7)

```

With N=7 the approximation of e is correct with four decimals.

### Question 6 
 Create a function called filter_my_vector() with the arguments x and geq. The function should take
 a vector x and set all values greater than or equal to geq to missing value (NA).


```{r}

filter_my_vector <- function(x,geq){
  
  x[x>=geq]<-NA
  return(x)
}


filter_my_vector(x = c(2, 9, 2, 4, 102), geq = 4)

```

### Question 7 
 create a magic matrix


```{r}

my_magic_matrix<- function (){
  
  x<-c(4,9,2,3,5,7,8,1,6)
  Mat<- matrix(x,nrow = 3,ncol = 3,byrow = TRUE)
  return(Mat)
  
}
my_magic_matrix()

# every row and column adds up to 15
```


### Question 8
Create a function called calculate_elements(A) that can take a matrix of an arbitrary size and    
calculate the number of elements in the matrix.


```{r}

Calculate_elements <- function(A){
  
  Result<- nrow(A)*ncol(A) 
  return(Result)
  
}

mat<-my_magic_matrix()

Calculate_elements(A=mat)

new_mat <- cbind(mat, mat)
Calculate_elements(A = new_mat)



```


### Uppgift 9

Create a function called row_to_zero(A, i) that can take a matrix of an arbitrary size and set the
row indexed with i to zero.


```{r}

row_to_zero<-function(A,i){
  
  A[i,]<-0
  return(A)
}


row_to_zero(A=mat,i=3)

row_to_zero(A=mat,i=1)

```

### Uppgift 10 
Create a function called add_elements_to_matrix() with parameters A, x, i, j. The function should
take a matrix A of an arbitrary size and add the value x to the parts of A indexed by row(s) i and
column(s) j.



```{r}

add_elements_to_matrix<-function(A,x,i,j){
  
  A[i,j]<- A[i,j]+x
  return(A)
  
}

add_elements_to_matrix(A = mat, x = 10, i = 2, j = 3)

add_elements_to_matrix(A = mat, x = -2, i = 1: 3, j = 2: 3)


```


### Question 11 


```{r}

my_magic_list<- function(){
  
my_text<-"My own list"
my_vector<-my_num_vector()
my_matrix<-my_magic_matrix()

listan<- list(info= my_text,my_vector,my_matrix)
return(listan)
}

my_magic_list()

```



### Question 12 
Create a function that will take a list x (that must contain one element with name info) and change
this element to the text argument given by text.

```{r}

change_info <- function(x, text) {
  x[1] <- text
  return(x)
}

change_info(x=my_magic_list(),text="some new info")



```
### Question 13 

Create a function called sum_numeric_parts() that will take a list x and sum together all numeric
elements in this list.

```{r}

sum_numeric_parts <-function(x){
  sum_num<- sum(as.numeric(unlist(x)),na.rm = TRUE)
  return(sum_num)
}

  
a_list<-my_magic_list()

sum_numeric_parts(x=a_list)
sum_numeric_parts(x=a_list[2])

```
### Question 14 

Create a function that will take a list x and add a new list element with the name note. This new element should contain text from the note parameter. 


```{r}


add_note <- function(x,note){
  New_list_element<- note
  old_list<- x
  ret<- list(info= old_list,note=New_list_element)
  return(ret)
}

a_list <- my_magic_list()
add_note(x = a_list, note = "This is a magic list!")


```
### Question 15 

Create a function that generates a data.frame


```{r}

my_data.frame<-function(){
 id <- c(1,2,3)
     name <- c("John", "Lisa", "Azra")
     income <- c(7.30,0.00, 15.21)
     rich <- c(FALSE, FALSE, TRUE)
     df <- data.frame(id = id, name = name, income = income, rich = rich)
  return(df)
  
}

my_data.frame()

```

### Question 16 

Sorting a data.frame by one variable 

```{r}

sort_head<- function(df,var.name,n){
  
  sorting<- head(sort(df[,var.name],decreasing=TRUE),n)
  return(sorting)
  
}

sort_head(df=iris,var.name="Petal.Length",n=5)

```
```{r}
# Alternativley 

sort_head2 <- function(df,var.name,n){
  sort2<-head(df[order(df[,var.name],decreasing = TRUE),],n)
  return(sort2)
  
}

sort_head2(df=iris,var.name="Petal.Length",n=12)


```

### Question 17 

Create a compared to median function 


```{r}

add_median_variable<-function(df,j){
  
  Median<- median(df[,j])
  
  df$Compared_to_median<- ifelse(df[,j]==Median,c("Median"),ifelse(df[,j]<Median,c("Smaller"),c("Greater")))
  
  return(df)
  
}

head(add_median_variable(df = faithful, 1), n = 12)


```

### Question 18 


```{r}


analyze_columns<- function(df,j){
  
  Mean<- apply(df[,j],2,mean)
  Median<- apply(df[,j],2,median)
  SD<- apply(df[,j],2, sd)
  
  Statistics <- rbind(Mean,Median,SD)
  
  Corr <- cor(df[,j])
  
  Listan<- list(Statistics,Correlation_Matrix=Corr)
  
  return(Listan)
  
}

analyze_columns(df = faithful, 1:2)
analyze_columns(df = iris, c(1,3))

# Alternatively for better formatting you can use lapply 
# It returns elements as list objects from vectors etc

analyze_columns <- function(df, j){
  
    mean <- data.frame(lapply(df[,j], mean))
    median <- data.frame(lapply(df[,j], median))
    sd <- data.frame(lapply(df[,j], sd))
    rb <- as.data.frame(rbind(mean, median, sd))
    cor <- cor(df[,j])
    out <- list(c(rb), correlation_matrix = cor)
    
    
    return(out)
    }
  
  
analyze_columns(df = faithful, 1:2)
analyze_columns(df = iris, c(1,3))
  



```
### Question 19 

Trace of a matrix 


```{r}

A <- matrix(2:5, nrow = 2)
B <- matrix(1:9, nrow = 3)
C <- matrix(9:-6, nrow = 4)

matrix_trace <- function (X){
  X[upper.tri(X)] <- 0
  X[lower.tri(X)] <- 0
  sum <- sum(X)
  return(sum)
}

matrix_trace(A)

# Alternatively 

matrix_trace1 <- function (X){
  sum_diag <- sum(diag(X))
  return(sum_diag)
}

matrix_trace1(A)

```






















