---
title: "Assignment 2"
author: "Edvin Magnusson"
date: ''
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Question 1 

Logical Equality 

```{r}

logical_equality <- function(A,B){
  
  A<- as.logical(A)
  B<- as.logical(B)
  
  AB<- A==B
  
  return(AB)
}

logical_equality(A = TRUE, B = FALSE)

logical_equality(A = FALSE, B = FALSE)
logical_equality(A = T, B = T)


```
### Question 2 

Sheldon game (Rock-paper-scissors-lizard-spock). 

```{r}

sheldon_game <- function(player1, player2) {
  choices <-
    c("Rock", "Paper", "Scissors", "Lizard", "Spock") # The valid schoices
  stopifnot(player1 %in% choices, player2 %in% choices) # Stop if not a valid choice
  choice1 <- which(choices %in% player1) # choice for player 1
  choice2 <- which(choices %in% player2) # choice for player 2
  if (any((choice1 == 1 &
           choice2 ==  (3 |
                        4)) |
          (choice1 == 2 &
           choice2 == (1 |
                       5)) | (choice1 == 3 & choice2 == (2 | 4)) | # The Sheldon rules
          (choice1 == 4 &
           choice2 == (5 | 2)) | (choice1 == 5 & choice2 == (1 | 3))
  )) {
    return("Player 1 wins!")
  } else if (choice1 == choice2) {
    return("Draw!")
  } else {
    return("Player 2 wins!")
  }
}

sheldon_game("Paper", "Rock")

sheldon_game("Spock", "Spock")

```
### Question 3 

Moving Median  

```{r}
#Using the ellipse ... to have unspecified variables 

my_moving_median <- function(x,n,...){
  
  if(!is.numeric(x)|!is.numeric(n)){
    stop("x or n is not numeric")
  }
  
  z<-c()
  i<-1
  while(i<=length(x)-n){
    z[i]<-median(x[seq(from =i,to=i+n)],...) # Here we include it to later be able to call the na.rm=T
    i<-i+1
  }
  return(z)
}


my_moving_median(x = 1:10, n = 2)

my_moving_median(x = 5:15, n = 4)

my_moving_median(x = c(5, 1, 2, NA, 2, 5, 6, 8, 9, 9), n = 2)

my_moving_median(x = c(5, 1, 2, NA, 2, 5, 6, 8, 9, 9), n = 2, na.rm = TRUE)


```


### Question 4 

Multiplication table

```{r}

for_mult_table<- function(from,to){
  
   if(!is.numeric(from)|!is.numeric(to)){
    stop("x or n is not numeric")
  }
  
Mat<- matrix(nrow =length(from:to),ncol = length(from:to) )

rows<-from:to
cols<-from:to
rownames(Mat)<-rows
colnames(Mat)<-cols

for(i in 1:dim(Mat)[1]){
  for(j in 1:dim(Mat)[2]){
    Mat[i,j]<- rows[i]*cols[j]
  }
  
}
  return(Mat)
  
}

for_mult_table(from = 1, to = 5)

for_mult_table(from = 10, to = 12)


```

### Question 5

Correlation matrix. 

```{r,error=TRUE}

cor_matrix<-function(x){
  
  if(!is.data.frame(x)){
    stop("Data is not a dataframe!!")
  }
  
  Mat<- scale(x,center = TRUE, scale = TRUE)
  n<-nrow(x)
  correlation_matrix<-(t(Mat)%*%Mat)/(n-1)

  return(correlation_matrix)  
  
}

data(iris)
cor_matrix(iris[, 1:4])

cor_matrix(as.list(iris[,1:4]))


```



### Question 6 

Calculating a cumulative sum and stopping when find_sum is traversed.

```{r}

find_cumsum<- function(x,find_sum){
  
  if(!is.numeric(x)|!is.numeric(find_sum)){
    stop("Arguments are nor numeric!")
  }
  
  
  sum1 <- 0 
  i<- 1
  
  while(i<=length(x)){
    
    sum1<-sum1+i
    i<-i+1
    if(sum1>find_sum)
      break;
    
  }
  
  return(sum1)
}

find_cumsum(x = 1:100, find_sum = 500)

find_cumsum(x = 1:10, find_sum = 500)



```
# Question 7 

Multiplication table as in question 4 but with a while loop. 
Tricky one where you have to think about the columns and rows. 

```{r}

while_mult_table<- function(from,to){
  
  if(!is.numeric(from)|!is.numeric(to)){
    stop("From and to is not numeric")
  }
  
  rows <- from:to
  cols<- from:to
  mat<- matrix(nrow = length(rows),ncol = length(cols)) # Empty
  rownames(mat)<-rows
  colnames(mat)<-cols
  
  i<--length(rows)
  j<--length(cols)
  while(i<= dim(mat)[1]){
    while(j<= dim(mat)[2]){
      
    mat[i,j]<-rows[i]*cols[j]
    j<-j+1
   i<-i
    }
   i<-i+1
   j<-i
    
  }
  return(mat)
  
}


while_mult_table(from = 3, to = 5)


while_mult_table(from = 7, to = 12)

```



### Question 8 


Same as question 6 but with a repeat loop. 


```{r}

repeat_find_cumsum<-function(x,find_sum){
  if(!is.numeric(x)|!is.numeric(find_sum)){
    stop("Arguments are not numeric")
  }
  
  sum1<- 0
  i<-1
  
  repeat{
    
    sum1<-sum1+i
    i<-i+1
    
    if(i>length(x))
      break()
    
    if (sum1>find_sum)
      break()
    
  }
  return(sum1)
  
}

repeat_find_cumsum(x = 1:100, find_sum = 500)

repeat_find_cumsum(x = 1:10, find_sum = 500)


```

### Question 9 

Moving median again but with a repeat loop. 

```{r}

repeat_my_moving_median<- function(x,n,...){
  if(!is.numeric(x)|!is.numeric(n)){
    stop("Arguments are not numeric!")
  }
  
  z<-c()
  i<-1 
  
  repeat{
    
    
    z[i]<- median(x[seq(from=i,to=i+n)],na.rm = TRUE)
    i<-i+1
    
    if (i>length(x))
      break()
  }
  
  return(z)
}


repeat_my_moving_median(x=1:10,n=2)

repeat_my_moving_median(x = 5:15, n = 4)

```

### Question 10 

lapply applies functions on list objects and returns a list object of the same length. 
A data frame is actually a type of list, so we can use lapply on a data frame without first converting it with as.list()

```{r}

coefvar <- function(X){
  
  if(!is.data.frame(X)){
    stop("X is not a dataframe")
  }
  z<- c(lapply(X,
               FUN = function(X)
                 sd(X)/mean(X)
               ))
  
  return(z)
}


data("iris")

coefvar(X=iris[1:4])
```

### Question 11 

Function for calculating bmi, the function gives a warning message if weight or height is <= 0.

```{r}

bmi<-function(body_weight,body_height){
  
  bmi<- body_weight/body_height^2
  
  if (body_weight<=0)
    warning("body_weight is not positive, calculation is not meaningful")
  
   if (body_height<=0)
    warning("body_heigth is not positive, calculation is not meaningful")
}

bmi(body_weight = 95, body_height = 1.98)

bmi(body_weight = 74, body_height = -1.83)

```

### Question 12 

Babylon method for approximating the square root of a number. 

```{r}

babylon <-  function(x, init, tol) { 
  sqrt_approx <- x 
  prop <- init  
  new <- (prop + x / prop) / 2  
  iter <- 0 
  while (abs(new - prop) > tol) {   
  iter <- iter + 1   
  prop <-  new   
  new <-  (prop + x / prop) / 2   
  rot <-  new   
  print(prop) 
  }
  return(list(Iterations = iter, Sqrt = rot))
}


babylon(40, 20, 0.1)


babylon(x = 2, init = 1.5, tol = 0.01)
```
### Question 13 

Hilbert matrix with nested for loop. 
 
 
```{r}

hilbert_matrix <- function(nrrow,nrcol){
  
  Mat<-matrix(nrow = nrrow,ncol = nrcol)
  
  for(i in 1:nrrow){
   for(j in 1:nrcol)
     Mat[i,j]<- 1/(i+j-1)
  }
 return(Mat) 
}

hilbert_matrix(nrrow=5, nrcol=2)

hilbert_matrix(nrrow=5, nrcol=4)
```


### Question 14 

Toeplitz matrix 


```{r,error=TRUE}

toeplitz_matrix <- function(x) {
  if (length(x) %% 2 == 0) {
    stop("Length of vector x is even")
  }
  res_mat <-matrix(nrow = ceiling(length(x) / 2), ncol = ceiling(length(x) / 2))
  n <-  ceiling(length(x) / 2)
  res_mat[1, ] <- x[1:n]
  res_mat[-1, 1] <- x[(n + 1):length(x)]
  for (i in 2:n) {
    for (j in 2:n) {
      res_mat[i, j] <- res_mat[i - 1, j - 1]
      
      
    }
    
    
  }
  return(res_mat)
}



toeplitz_matrix(x = 1:5)
toeplitz_matrix(x = 1:4)

toeplitz_matrix(c(1, 0, 2, 0, 3, 0, 4, 0, 5))

```








